---
title: "Activity Analysis"
author: "Beth Hastings"
date: "July 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This report analyzes activity data to demonstrate knowledge of R Markdown  for the Reproducible
Research class assignment.


## Loading the Data

```{r load_libraries}
library(dplyr)
library(ggplot2)
library(mosaic)
library(xtable)
library(knitr)
```
1. Code for reading in the dataset and/or processing the data


```{r load_data}
d<- read.csv(file="c:/Users/bh17649/Documents/Data_Science/Course Notes/Reproducible Research/Class Projects/activity.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)
names(d)
head(d)
nrow(d)
ncol(d)
```

2. Histogram of the total number of steps taken each day

```{r histogram}
#Calculate by day
s<- d %>% group_by(date) %>% summarize(total_steps = sum(steps))
p<- ggplot(s, aes(total_steps))

p + geom_histogram(binwidth = 2000) + xlab("Steps") + ylab("Observations") + labs(title = "Histogram of Step Observations")
```

3. Mean and median number of steps taken each day
```{r mean_med_steps_per_day}
#Check Mean, Median and Quartile results by day
summ<- s %>% summarize(mean_steps = mean(total_steps, na.rm=TRUE), median_steps=median(total_steps, na.rm=TRUE))
summ

```
4. Time series plot of the average number of steps taken
```{r time_series_by_day}
d<- as.data.frame(d)
d$date2<- as.character(d$date)
d$date2<- as.Date(d$date)

ggplot(s, aes(date, total_steps, group=1)) +
           geom_point() +
           geom_line() +  
           ggtitle("Steps by Day") +
           xlab("Date") + ylab("Total Steps")  +
           theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 20)) +
           theme(axis.text.x = element_text(angle=60, hjust=1))

```


5. The 5-minute interval that, on average, contains the maximum number of steps
```{r max_interval}
i<- d %>% select(interval, steps) %>% group_by(interval) %>% summarize(avg_steps = mean(steps, na.rm = T))

i<- arrange(i, desc(avg_steps))

kable(head(i))
```

6. Code to describe and show a strategy for imputing missing data

```{r impute_Val}
#Replace missing values with average for the interval
d$steps2<- d$steps

i<- d %>% 
        group_by(interval) %>% 
        mutate(steps2 = replace(steps2, is.na(steps), mean(steps, na.rm=TRUE)))

#Checck original and new variable to ensure all missing values are gone.
kable(favstats(i$steps))
kable(favstats(i$steps2))
```

7. Histogram of the total number of steps taken each day after missing values are imputed
```{r na.rm_histogram}

s<- i %>% group_by(date) %>% summarize(total_steps = sum(steps2))
p<- ggplot(s, aes(total_steps))

p + geom_histogram(binwidth = 2000) + xlab("Steps") + ylab("Observations") + labs(title = "Histogram of Step Observations")
```
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r weekend_weekday}
i$dayofweek <- weekdays(as.Date(d$date))
i$weekend_flg<-i$dayofweek
i<- i %>%  
     mutate(weekend_flg = ifelse((dayofweek == "Sunday" | dayofweek == "Saturday"), "Weekend",  "Weekday"))

i<- i%>% select(weekend_flg, interval, steps2) %>%
        group_by(interval, weekend_flg) %>%
        summarize(avg_steps= mean(steps2))

p<- ggplot(i, aes(interval, avg_steps, group=1))

p+         geom_line() +  
           ggtitle("Average Steps by Interval, Weekday vs. Weekend") +
           xlab("Interval") + ylab("Avg Steps")  +
           theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 20)) +
           theme(axis.text.x = element_text(angle=60, hjust=1)) +
           facet_wrap(~ i$weekend_flg, ncol=1)
```

9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

#See above :)!